# Background and use

This script is used to 1) import all variables used in this analysis using SQL, 2) clean and process the data for subsequent analyses.

# Setup

Install and load in necessary packages:

```{r}
install.packages("pacman")
pacman::p_load(
  tidyverse,
  bigrquery,
  jsonlite,
  mice,
  table1
)
```

Define environmental variables:

```{r}
library(jsonlite)

wb <- function(...) fromJSON(system(paste("wb", ..., "--format=json"), intern = TRUE))

workspace <- wb("workspace describe")
GOOGLE_CLOUD_PROJECT <- workspace$googleProjectId

resources <- wb("resource list")

# WORKSPACE_BUCKET
bucket_row <- resources[resources$resourceType == "GCS_BUCKET" &
                        grepl("workspace-bucket", resources$id) &
                        !grepl("temporary", resources$id), ]
WORKSPACE_BUCKET <- paste0("gs://", bucket_row$bucketName[1])

# WORKSPACE_CDR
bq <- resources[resources$resourceType %in% c("BQ_DATASET", "BIGQUERY_DATASET"), ]
cdr_row <- bq[grepl("^C\\d{4}Q\\d+R\\d+$", bq$datasetId), ]
WORKSPACE_CDR <- paste0(cdr_row$projectId[1], ".", cdr_row$datasetId[1])

# Summary
for (var in c("GOOGLE_CLOUD_PROJECT", "WORKSPACE_BUCKET", "WORKSPACE_CDR")) {
  val <- get(var)
  cat(var, ":", if (length(val) && nzchar(val)) val else "NOT FOUND", "\n")
}
```

Define necessary function:

```{r}
download_data <- function(query) {
  tb <- bq_project_query(GOOGLE_CLOUD_PROJECT, 
                         query = str_glue(query), 
                         default_dataset = WORKSPACE_CDR)
  bq_table_download(tb)
}
```

# Data import

## Covariates

All covariates are imported in the same way:

1. Import raw variable data using SQL
2. Minimally process the variable
3. Merge into one covariate dataframe

### EHR

```{r}
ehr_query <- "
    SELECT
        person_id,
        MIN(condition_start_date) AS date_first_ehr,
        MAX(condition_start_date) AS date_last_ehr
    FROM
        condition_occurrence
    GROUP BY
        person_id
"

df_ehr <- download_data(ehr_query)

df_ehr <- df_ehr %>%
    mutate(length_ehr = as.numeric(date_last_ehr - date_first_ehr))
```

### Sex at birth

```{r}
sex_query <- "
    SELECT
        person_id,
        answer AS sex
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1585845))
"

df_sex <- download_data(sex_query)

df_sex <- df_sex %>%
    mutate(sex = case_when(
        sex == "Sex At Birth: Female" ~ 1,
        sex == "Sex At Birth: Male" ~ 2,
        sex == "Sex At Birth: Intersex" ~ NA_integer_,
        sex == "Sex At Birth: Sex At Birth None of These" ~ NA_integer_,
        sex == "PMI: Skip" ~ NA_integer_,
        sex == "PMI: Prefer Not To Answer" ~ NA_integer_),
           sex = factor(sex,
                        levels = c(1, 2),
                        labels = c("Female", "Male")))

df_covar <- df_ehr %>%
    left_join(df_sex, by = "person_id")
```

### Race/ethnicity

```{r}
race_eth_query <- "
    SELECT
        person_id,
        answer AS race_eth
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1586140))
"

df_race_eth <- download_data(race_eth_query)

df_race_eth <- df_race_eth %>%
    mutate(
        race_eth_code = case_when(
            race_eth == "What Race Ethnicity: White" ~ 1,
            race_eth == "What Race Ethnicity: Black" ~ 2,
            race_eth == "What Race Ethnicity: Hispanic" ~ 3,
            race_eth %in% c("PMI: Skip", "PMI: Prefer Not To Answer") ~ NA_integer_,
            TRUE ~ 4
        )
    )

df_race_eth <- df_race_eth %>%
    group_by(person_id) %>%
    summarise(
        race_eth_code = if (n_distinct(na.omit(race_eth_code)) > 1) 4 else first(na.omit(race_eth_code)),
            .groups = "drop"
        ) %>%
    mutate(
        race_eth = factor(race_eth_code,
                          levels = c(1, 2, 3, 4),
                          labels = c("White", "Black or African American", "Hispanic", "Other"))
  ) %>%
    select(-race_eth_code)

df_covar <- df_covar %>%
    left_join(df_race_eth, by = "person_id")
```

### Income

```{r}
income_query <- "
    SELECT
        person_id,
        answer AS income
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1585375))
"

df_income <- download_data(income_query)

df_income <- df_income %>%
    mutate(income = case_when(
        income == "Annual Income: less 10k" ~ 1,
        income == "Annual Income: 10k 25k" ~ 1,
        income == "Annual Income: 25k 35k" ~ 1,
        income == "Annual Income: 35k 50k" ~ 2,
        income == "Annual Income: 50k 75k" ~ 2,
        income == "Annual Income: 75k 100k" ~ 3,
        income == "Annual Income: 100k 150k" ~ 3,
        income == "Annual Income: 150k 200k" ~ 4,
        income == "Annual Income: more 200k" ~ 4,
        income == "PMI: Skip" ~ 5,
        income == "PMI: Prefer Not To Answer" ~ 5),
          income = factor(income,
                          levels = c(1, 2, 3, 4, 5),
                          labels = c("<35k", "35-75k", "75-150k", ">150k", "Prefer not to answer")))

df_covar <- df_covar %>%
    left_join(df_income, by = "person_id")
```

### Alcohol

```{r}
alcohol_ever_query <- "
    SELECT
        person_id,
        answer AS alcohol_ever
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1586198))
"

df_alcohol_ever <- download_data(alcohol_ever_query)

alcohol_freq_query <- "
    SELECT
        person_id,
        answer AS alcohol_freq
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1586201))
"

df_alcohol_freq <- download_data(alcohol_freq_query)

df_alcohol <- df_alcohol_ever %>%
    full_join(df_alcohol_freq, by = "person_id")

df_alcohol <- df_alcohol %>%
    mutate(alcohol = case_when(
        alcohol_ever == "Alcohol Participant: No" ~ 1,
        alcohol_ever == "Alcohol Participant: Yes" & alcohol_freq == "Drink Frequency Past Year: Never" ~ 2,
        alcohol_freq == "Drink Frequency Past Year: Monthly Or Less" ~ 3,
        alcohol_freq == "Drink Frequency Past Year: 2 to 4 Per Month" ~ 3,
        alcohol_freq == "Drink Frequency Past Year: 2 to 3 Per Week" ~ 4,
        alcohol_freq == "Drink Frequency Past Year: 4 or More Per Week" ~ 4,
        alcohol_ever == "PMI: Prefer Not To Answer" & alcohol_freq == "PMI: Prefer Not To Answer" ~ NA_integer_,
        alcohol_ever == "PMI: Skip" & alcohol_freq == "PMI: Skip" ~ NA_integer_),
           alcohol = factor(alcohol,
                   levels = c(1, 2, 3, 4),
                   labels = c("Never alcohol consumer",
                              "Past alcohol consumer",
                              "Current alcohol consumer - monthly", 
                              "Current alcohol consumer - weekly"))) %>%
    select(-c("alcohol_ever", "alcohol_freq"))

df_covar <- df_covar %>%
    left_join(df_alcohol, by = "person_id")
```

### Smoking

```{r}
smoking_query <- "
    SELECT
        person_id,
        answer AS smoking
    FROM
        ds_survey
    WHERE
        (question_concept_id IN (1585857))
"

df_smoking <- download_data(smoking_query)

df_smoking <- df_smoking %>%
    mutate(smoking = case_when(
        smoking == "100 Cigs Lifetime: Yes" ~ 1,
        smoking == "100 Cigs Lifetime: No" ~ 2,
        smoking == "PMI: Dont Know" ~ NA_integer_,
        smoking == "PMI: Prefer Not To Answer" ~ NA_integer_,
        smoking == "PMI: Skip" ~ NA_integer_),
           smoking = factor(smoking,
                   levels = c(1, 2),
                   labels = c("Current or past smoker",
                              "Never smoker")))

df_covar <- df_covar %>%
    left_join(df_smoking, by = "person_id")
```

### BMI

```{r}
height_query <- "
    SELECT
        person_id,
        value_as_number AS height,
        unit_source_value AS height_unit,
        measurement_datetime AS height_datetime,
        measurement_date
    FROM
        measurement
    WHERE
        (measurement_source_concept_id IN (903133))
"

df_height <- download_data(height_query)

df_height <- df_height %>%
    group_by(person_id) %>%
    slice_min(height_datetime, with_ties = FALSE) %>% # Get first height measurement
    ungroup() %>%
    na.omit()

weight_query <- "
    SELECT
        person_id,
        value_as_number AS weight,
        unit_source_value AS weight_unit,
        measurement_datetime AS weight_datetime,
        measurement_date
    FROM
        measurement
    WHERE
        (measurement_source_concept_id IN (903121))
"

df_weight <- download_data(weight_query)

df_weight <- df_weight %>%
    group_by(person_id) %>%
    slice_min(weight_datetime, with_ties = FALSE) %>% # Get first weight measurement
    ungroup() %>%
    na.omit()

df_bmi <- df_height %>%
    inner_join(df_weight, by = c("person_id", "measurement_date")) %>%
    mutate(
        bmi = weight / ((height / 100)^2)
    )

df_bmi <- df_bmi %>%
    mutate(bmi_cat = case_when(
        bmi < 18.5 ~ 1,
        bmi >= 18.5 & bmi < 25 ~ 2,
        bmi >= 25 & bmi < 30 ~ 3,
        bmi >= 30 ~ 4,
        TRUE ~ NA_integer_),
           bmi_cat = factor(bmi_cat,
                   levels = c(1, 2, 3, 4),
                   labels = c("Underweight",
                              "Healthy weight",
                              "Overweight",
                              "Obesity"))) %>%
    select(person_id, bmi_cat)

df_covar <- df_covar %>%
    left_join(df_bmi, by = "person_id")
```

### Age

```{r}
age_query <- "
    SELECT
        person_id,
        CAST(person.birth_datetime AS DATE) AS date_of_birth
    FROM
        person
"

df_age <- download_data(age_query)

df_covar <- df_covar %>%
    left_join(df_age, by = "person_id")

df_covar <- df_covar %>%
    mutate(age_last_aou = time_length(interval(date_of_birth, as.Date("2023-10-01")), "years"))
```

### Imputation

```{r}
data.frame(
  NA_Count = colSums(is.na(df_covar)),
  Proportion = colMeans(is.na(df_covar))
)
```

Use regression imputation on missing values:

*Note that sex is not imputed as it is a necessary variable for subsequent analyses*

```{r}
df_covar_impute <- df_covar %>% filter(!is.na(sex))

vars_to_impute <- c("race_eth", "income", "alcohol", "smoking", "bmi_cat")
vars_for_model <- c("sex", vars_to_impute)

df_mice <- df_covar_impute %>% select(all_of(vars_for_model))
original_order <- names(df_covar_impute)

method <- make.method(df_mice)
method["sex"]        <- ""
method["race_eth"]   <- "polyreg"
method["income"]     <- "polyreg"
method["alcohol"]    <- "polyreg"
method["smoking"]    <- "logreg"
method["bmi_cat"]    <- "polyreg"

imp <- mice(df_mice, m = 1, method = method, seed = 123)

df_imputed <- complete(imp, 1)

df_output <- df_covar_impute
df_output[vars_to_impute] <- df_imputed[vars_to_impute]

df_covar <- df_output[, original_order]
```

```{r}
data.frame(
  NA_Count = colSums(is.na(df_covar)),
  Proportion = colMeans(is.na(df_covar))
)
```

### Summarize the covariate data

```{r}
label(df_covar$age_last_aou) <- "Age at All of Us censor date"
label(df_covar$sex) <- "Sex"
label(df_covar$race_eth) <- "Race/ethnicity"
label(df_covar$alcohol) <- "Alcohol consumption"
label(df_covar$smoking) <- "Smoking history"
label(df_covar$income) <- "Annual income"
label(df_covar$bmi_cat) <- "Body mass index"

output <- table1(~ age_last_aou + sex + race_eth + alcohol + smoking + income + bmi_cat, 
                 data = df_covar, 
                 render.continuous=c("Median [IQR]"="Median [Q1, Q3]"))

tab1 <- as.data.frame(output)
head(tab1, n = 40)
```

Save the dataframe:

```{r}
write_csv(df_covar, file = "df_covar.csv")

system2("gcloud",
        args = c("storage", "cp", "df_covar.csv", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
```

## Fitbit

### Step count (steps)

```{r}
steps_query <- "
    SELECT
        person_id,
        DATE(datetime) AS date,
        SUM(steps) AS steps
    FROM
        steps_intraday
    WHERE
        steps > 0
    GROUP BY
        person_id,
        DATE(datetime);
"

df_steps <- download_data(steps_query)
```

### Peak 1-min cadence (cadence1)

```{r}
cadence1_query <- "
    SELECT
        person_id,
        DATE(datetime) AS date,
        MAX(steps) AS cadence1
    FROM
        steps_intraday
    WHERE
        steps > 0
    GROUP BY
        person_id,
        DATE(datetime);
"

df_cadence1 <- download_data(cadence1_query)
```

```{r}
df_exposure <- df_steps %>%
    merge(df_cadence1, by = c("person_id", "date"))
```

### Peak 30-min cadence (cadence30)

```{r}
cadence30_query <- "
    WITH ranked_steps AS (
        SELECT
            person_id,
            DATE(datetime) AS date,
            steps,
            ROW_NUMBER() OVER (
                PARTITION BY person_id, DATE(datetime)
                ORDER BY steps DESC
            ) AS step_rank
        FROM steps_intraday
    )
    SELECT
        person_id,
        date,
        AVG(steps) AS cadence30
    FROM
        ranked_steps
    WHERE
        step_rank <= 30
    GROUP BY
        person_id,
        date;
"

df_cadence30 <- download_data(cadence30_query)
```

```{r}
df_exposure <- df_exposure %>%
    merge(df_cadence30, by = c("person_id", "date"))
```

### Daily heart rate per step (dhrps)

```{r}
dhrps_query <- "
    SELECT
        person_id,
        DATE(datetime) AS date,
        AVG(heart_rate_value) AS avg_hr
    FROM
        heart_rate_minute_level
    GROUP BY
        person_id,
        DATE(datetime);
"

df_dhrps <- download_data(dhrps_query)
```

```{r}
df_exposure <- df_exposure %>%
    merge(df_dhrps, by = c("person_id", "date"))
```

```{r}
df_exposure <- df_exposure %>%
    mutate(dhrps = avg_hr / steps)
```

### WEAR/BYOD

```{r}
wear_query <- "
    SELECT
        person_id,
        resultsconsent_wear AS wear
    FROM
        wear_study
"

df_wear <- download_data(wear_query)
```

```{r}
df_exposure <- df_exposure %>%
    left_join(df_wear, by = "person_id") %>%
    mutate(wear_status = factor(ifelse(!is.na(wear) & wear == "Yes", 1, 2),
                                levels = c(1, 2),
                                labels = c("WEAR", "BYOD"))) %>%
    select(-wear)
```

### Wear time

```{r}
wear_time_query <- "
    SELECT
        person_id,
        DATE(datetime) AS date,
        COUNT(DISTINCT FORMAT_TIMESTAMP('%Y-%m-%d %H:%M', datetime)) AS wear_time_minutes
    FROM
        heart_rate_minute_level
    GROUP BY
        person_id,
        DATE(datetime)
"
df_wear_time <- download_data(wear_time_query)
```

```{r}
df_exposure <- df_exposure %>%
    merge(df_wear_time, by = c("person_id", "date"))
```

Save the dataframe:

```{r}
write_csv(df_exposure, file = "df_exposure.csv")

system2("gcloud",
        args = c("storage", "cp", "df_exposure.csv", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
```

# Data cleaning

## Exclusions

Initialize a table to keep track of exclusions:

```{r}
table_exclusions <- data.frame("Cohort" = "Cohort with EHR data",
                               "Participants" = n_distinct(df_covar$person_id),
                               "Participants_excluded" = NA,
                               "Observations" = nrow(df_covar),
                               "Observations_excluded" = NA)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "Cohort with Fitbit data", 
                                     "Participants" = n_distinct(df_exposure$person_id),
                                     "Participants_excluded" = NA,
                                     "Observations" = nrow(df_exposure),
                                     "Observations_excluded" = NA))
```

Participant is excluded if:
1. They don't have both covariate and Fitbit data

Day of data is excluded if:
2. Age <18
3. Step count <100
4. Step count >75000
5. Peak 1-min or peak 30-min cadence >210
6. Average HR <20
7. Average HR >220
8. Daily heart rate per step <3 standard deviations below the mean
9. Daily heart rate per step >3 standard deviations above the mean
10. Wear time <16 hours

```{r}
# 1 (data availability exclusion)
participants <- n_distinct(c(df_exposure$person_id, df_covar$person_id))
observations <- nrow(df_exposure) + nrow(df_covar)

df <- df_covar %>%
    merge(df_exposure, by = "person_id")

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "Cohort with EHR and Fitbit data", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
# 2 (age exclusion)
df <- df %>%
    mutate(age_at_date = as.numeric(floor((date - date_of_birth)/365.25)))

participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(age_at_date >= 18)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All age >= 18", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
# 3 and 4 (step count exclusion)
participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(steps >= 100 & steps <= 75000)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All step count >=100 and <= 75000 steps", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
# 5 (cadence exclusion)
participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(cadence1 <= 210)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All cadence1 <= 210 steps/min", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))

participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(cadence30 <= 210)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All cadence30 <= 210 steps/min", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))

```

```{r}
# 6 and 7 (heart rate exclusion)
participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(avg_hr >= 20 & avg_hr <= 220)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All avg_hr >= 20 and <= 220 bpm", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
# 8 and 9 (dhrps exclusion)
three_sd_above_mean = mean(df$dhrps) + 3 * sd(df$dhrps)
three_sd_below_mean = mean(df$dhrps) - 3 * sd(df$dhrps)

participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(dhrps >= three_sd_below_mean & dhrps <= three_sd_above_mean)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All DHRPS within three sd of mean", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
# 10 (wear time exclusion)
participants <- n_distinct(df$person_id)
observations <- nrow(df)

df <- df %>%
    filter(wear_time_minutes >= 16*60)

table_exclusions <- rbind(table_exclusions, 
                          data.frame("Cohort" = "All wear count >= 16 hours", 
                                     "Participants" = n_distinct(df$person_id),
                                     "Participants_excluded" = participants - n_distinct(df$person_id),
                                     "Observations" = nrow(df),
                                     "Observations_excluded" = observations - nrow(df)))
```

```{r}
table_exclusions
```

## Processing

Fitbit timing variables:

```{r}
df <- df %>%
    group_by(person_id) %>%
    mutate(
        date_first_fitbit = min(date, na.rm = TRUE),
        date_last_fitbit = max(date, na.rm = TRUE),
        fitbit_duration = as.numeric(date_last_fitbit - date_first_fitbit)
    ) %>%
    ungroup()
```

Time variables:

```{r}
df <- df %>%
    mutate(
        date_t0 = date_first_fitbit,
        date_tw = date_first_fitbit + days(6),
        date_tm = date_first_fitbit + days(29),
        date_ts = date_first_fitbit + days(179),
        date_ty = date_first_fitbit + days(364),
        date_t0_washout = date_t0 + days(14),
        date_tw_washout = date_tw + days(14),
        date_tm_washout = date_tm + days(14),
        date_ts_washout = date_ts + days(14),
        date_ty_washout = date_ty + days(14)
    )
```

Save as per person dataframe:

```{r}
df_per_pid <- df %>%
    group_by(person_id) %>%
    summarise(
        # covariates
        date_of_birth = first(date_of_birth),
        sex = first(sex),
        race_eth = first(race_eth),
        alcohol = first(alcohol),
        smoking = first(smoking),
        income = first(income),
        bmi_cat = first(bmi_cat),
        wear_status = first(wear_status),
        
        # timing variables
        date_first_fitbit = first(date_first_fitbit),
        date_last_fitbit = first(date_last_fitbit),
        fitbit_duration = first(fitbit_duration),
        date_first_ehr = first(date_first_ehr),
        date_last_ehr = first(date_last_ehr),
        date_last_aou = as.Date("2023-10-01"),
        date_t0 = first(date_t0),
        date_tw = first(date_tw),
        date_tm = first(date_tm),
        date_ts = first(date_ts),
        date_ty = first(date_ty),
        date_t0_washout = first(date_t0_washout),
        date_tw_washout = first(date_tw_washout),
        date_tm_washout = first(date_tm_washout),
        date_ts_washout = first(date_ts_washout),
        date_ty_washout = first(date_ty_washout),
        age_last_aou = first(age_last_aou)
    ) %>%
    ungroup()
```

```{r}
df_per_pid <- df_per_pid %>%
    mutate(
        fitbit_week = ifelse(fitbit_duration >= 7, 1, 0),
        fitbit_month = ifelse(fitbit_duration >= 30, 1, 0),
        fitbit_six = ifelse(fitbit_duration >= 180, 1, 0),
        fitbit_year = ifelse(fitbit_duration >= 365, 1, 0),
        age_t0 = as.numeric(date_t0 - date_of_birth) / 365.25
    )
```

## Phenotypes

```{r}
metrics <- c("steps", "cadence1", "cadence30", "dhrps")

time_periods <- list(
    day = list(
        filter_fn = function(df) filter(df, date == date_t0),
        prefix = "day"
    ),
    day_washout = list(
        filter_fn = function(df) filter(df, date == date_t0_washout),
        prefix = "day",
        suffix = "_washout"
    ),
    week = list(
        filter_fn = function(df) filter(df, date >= date_t0 & date < date_tw),
        prefix = "week"
    ),
    week_washout = list(
        filter_fn = function(df) filter(df, date >= date_t0_washout & date < date_tw_washout),
        prefix = "week",
        suffix = "_washout"
    ),
    month = list(
        filter_fn = function(df) filter(df, date >= date_t0 & date < date_tm),
        prefix = "month"
    ),
    month_washout = list(
        filter_fn = function(df) filter(df, date >= date_t0_washout & date < date_tm_washout),
        prefix = "month",
        suffix = "_washout"
    ),
    six = list(
        filter_fn = function(df) filter(df, date >= date_t0 & date < date_ts),
        prefix = "six"
    ),
    six_washout = list(
        filter_fn = function(df) filter(df, date >= date_t0_washout & date < date_ts_washout),
        prefix = "six",
        suffix = "_washout"
    ),
    year = list(
        filter_fn = function(df) filter(df, date >= date_t0 & date < date_ty),
        prefix = "year"
    ),
    year_washout = list(
        filter_fn = function(df) filter(df, date >= date_t0_washout & date < date_ty_washout),
        prefix = "year",
        suffix = "_washout"
    )
)

for (period_name in names(time_periods)) {
    period <- time_periods[[period_name]]
    suffix <- if (!is.null(period$suffix)) period$suffix else ""
  
    for (metric in metrics) {
        col_name <- paste0(period$prefix, "_", metric, suffix)
        q_name <- paste0(col_name, "_q")
        
        temp_df <- df %>%
        group_by(person_id) %>%
        period$filter_fn() %>%
        summarise(!!col_name := mean(.data[[metric]], na.rm = TRUE), .groups = "drop") %>%
        mutate(!!q_name := ntile(.data[[col_name]], 4))

        df_per_pid <- df_per_pid %>% left_join(temp_df, by = "person_id")
    }
}
```

Save to the bucket:

```{r}
write_csv(df, file = "df.csv")
saveRDS(df, "df.rds")

system2("gcloud",
        args = c("storage", "cp", "df.csv", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
system2("gcloud",
        args = c("storage", "cp", "df.rds", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)

write_csv(df_per_pid, file = "df_per_pid.csv")
saveRDS(df_per_pid, "df_per_pid.rds")

system2("gcloud",
        args = c("storage", "cp", "df_per_pid.csv", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
system2("gcloud",
        args = c("storage", "cp", "df_per_pid.rds", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
```

Save a version of the per PID dataframe for use in pheTK (all covariates must be saved as numeric variables):

```{r}
df_per_pid_phetk <- df_per_pid %>%
  mutate(across(
    c(sex, race_eth, alcohol, smoking, income, wear_status, bmi_cat),
    ~ as.integer(as.factor(.))
  ))

write_excel_csv(df_per_pid_phetk, 'df_per_pid_phetk.csv')
system2("gcloud",
        args = c("storage", "cp", "df_per_pid_phetk.csv", paste0(WORKSPACE_BUCKET, "/dataframes/")),
        stdout = TRUE,
        stderr = TRUE)
```
